Cosmic web alignments with the shape, angular momentum and peculiar velocities of dark matter halos

### 2014-03-24 paper/main.tex
typo

We study the alignment of dark matter halos with the cosmic web
characterized by the tidal and velocity shear fields. We focus on the
alignment of their shape, angular momentum and peculiar velocities. We
use a cosmological N-body simulation that allows to study dark matter
halos spanning almost five orders of magnitude in mass
($10^{9}$-$10^{14}$) \hMsun and use spatial scales of
$(0.5$-$1.0)$\hMpc to define the cosmic web. After varying the
numerical parameters in our experiments we find that the halo shape
presents the strongest alignment along filaments and walls defined by
the tidal field with a signal that gets stronger as the halo mass
increases; in the case of the velocity shear field only massive halos
$>10^{12}$\hMsun tend to have their shapes perpendicular to filaments
and walls. For the angular momentum we find alignment signals only for
halos more massive than $10^{12}$\hMsun both in the tidal and velocity
shear webs where the preferences are for it to be perpendicular to filaments
or parallel to walls, respectively. Finally, the peculiar velocities
show a strong alignment for all halo masses to move along filaments
and walls in the tidal web, while in the velocity shear the same
alignment is present but weaker and only for halos less massive than
$10^{12}$\hMsun. Most of the different signals between the two web
descriptions are due to the fact that the main directions of stronger
tidal tension and velocity shear coincide for halos below $10^{12}$\hMpc but are
perpendicular to each other for halos more massive than this
threshold.  Our results clearly show that the two different
algorithms we used to define the cosmic web describe different physical
aspects of non-linear collapse and should be used in a complementary
way to understand the effect of the cosmic web on galaxy evolution.
\end{abstract}
\begin{keywords}
methods: N-body simulations, galaxies: haloes, cosmology: theory, dark
matter, large-scale structure of Universe
\end{keywords}

>>>>

We study the alignment of dark matter halos with the cosmic web
characterized by the tidal and velocity shear fields. We focus on the
alignment of their shape, angular momentum and peculiar velocities. We
use a cosmological N-body simulation that allows to study dark matter
halos spanning almost five orders of magnitude in mass
($10^{9}$-$10^{14}$) \hMsun and use spatial scales of
$(0.5$-$1.0)$\hMpc to define the cosmic web. After varying the
numerical parameters in our experiments we find that the halo shape
presents the strongest alignment along filaments and walls defined by
the tidal field with a signal that gets stronger as the halo mass
increases; in the case of the velocity shear field only massive halos
$>10^{12}$\hMsun tend to have their shapes perpendicular to filaments
and walls. For the angular momentum we find alignment signals only for
halos more massive than $10^{12}$\hMsun both in the tidal and velocity
shear webs where the preferences are for it to be perpendicular to filaments
or parallel to walls, respectively. Finally, the peculiar velocities
show a strong alignment for all halo masses to move along filaments
and walls in the tidal web, while in the velocity shear the same
alignment is present but weaker and only for halos less massive than
$10^{12}$\hMsun. Most of the different signals between the two web
descriptions are due to the fact that the main directions of stronger
tidal tension and velocity shear coincide for halos below $10^{12}$\hMsun but are
perpendicular to each other for halos more massive than this
threshold.  Our results clearly show that the two different
algorithms we used to define the cosmic web describe different physical
aspects of non-linear collapse and should be used in a complementary
way to understand the effect of the cosmic web on galaxy evolution.
\end{abstract}
\begin{keywords}
methods: N-body simulations, galaxies: haloes, cosmology: theory, dark
matter, large-scale structure of Universe
\end{keywords}
